DTOADQ

### 2017-11-29 testdocs/_region_.tex
doicumentation

  In order to compute a normal of an SDF, you must perform a gradient
approximation: the central distance onf the SDF at P. For example, the
difference quotient:
  \[\frac{(f(x+h) - f(x))}{h}\]
  The most obvious way to calculate this is the six-point gradient; where for
each axis you compute $f(x+h) - f(x-h)$, and normalize the result. This
returns a vector pointing in the direction where the map's SDF changes the
most: the normal. A four point gradient is possible; as long as each each
point of the gradient is multiplied by $\epsilon$ before normalizing. The
speedup from six to four is worthwhile, as computing each gradient is
expensive as it involves a sdf map call. There are more utilities with
gradients not yet currently provided in DTOADQ but are worth mentioning, a
few are: an additional normal call using an alternative (higher-quality) SDF
map can be used to compute a 'normal' bumpmap, and analyzing the gradient by
hand allows a cheap normal computation of primitives such as a spheres.
  \section{OpenCL}
	OpenCL, and GPU programming in general, is where I am least knowledgeable or
interested. Well, OpenCL is easy to work with, and I wrote my own OpenCL module
to handle OCL in a rather convenient method, for example, the DTOADQ kernel call
looks as such:
	\begin{lstlisting}
	shared_info.image.Lock(); // Acquire OCL Mem from OGL image object
	ocl.Run(
	  ocl.CLStoreMem(shared_info.rw_image),
	  shared_info.image.RWrite(),
	  ocl.CLStoreMem(shared_info.image_metadata),
	  shared_info.camera,
	  shared_info.timer,
	  ocl.CLPredefinedMem(imgs),
	  ocl.CLStoreMem(shared_info.ocl_material),
	  debug_values,
	  ocl.CLStoreMem(shared_info.rng_states),
	  // kernel size
	  shared_info.image.x, shared_info.image.y
	);
	shared_info.image.Unlock();

@@

  The connection strategy below describes how to connect the edges of the two
paths. There is no $t = 0$ case as no physical camera lens exists on the scene.
  \begin{align}
    c_{s, t} =
    \begin{cases}
      L_e(L0 \rightarrow L1), &s = 0, t > 1,\\
      f_s(L0 \rightarrow L1 \rightarrow L2) G(E1 \leftrightarrow
      L1) f_s(L1 \rightarrow E1 \rightarrow E0), &s > 0, t > 1
    \end{cases}
  \end{align}
  In code:
  \begin{lstlisting}
Spectrum contribution = E1.irradiance * L1->irradiance;

@@

    Another question to consider is if the path should be able to reflect on the
light source. While under real circumstances, this is allowed, the contribution
is negligible and dismissed in DTOADQ. Also, there are no cases for $s = s, t =
0|1$ as the lens does not have a physical existence in the SDF map. The $t = 0$
case, where a photon from the light source hits the lens directly, is rather
pointless anyways, as a path generated by light-tracing would have to be able to
contribute to any pixel, which is not feasible on a GPU.
\\
    Moving on, for $\alpha^{L|E}_i$, these are equivalent for both $L$ and $E$,
the problem is that the $\bar{y}$ path is generated in perspective of the light
source, so there has to be a bit of caution in terms of perspective in when
generating the PDFs. $p_A$ is the PDF w.r.t. area, for the light path; we only
need to take the PDF of the solid angle of a sphere
  \begin{align}
    p_A^L(y_0 \rightarrow y_1) = p_{\sigma}^L(y_0 \rightarrow y_1) G(v
    \leftrightarrow \dot{v})\\
    p_{\sigma}^L(y_0 \rightarrow y_1) = \frac{1.0f}{\tau * (1.0f -
    \sqrt{\frac{R^2}{|y_0 - y_1|^2}})}
  \end{align}
    where $R$ is the radius of the sphere, note $y_0$ is a point on the surface
    of the sphere, and not its origin. For the eye:
  \begin{align}
    P_A^E = 1.0f
  \end{align}
    The $L^0_e$ is contribution from the light source. As of right now, DTOADQ
only supports area lights, so $L^0_e$ is set a constant for the emitter's
spectrum value. The $W_e^0$ is contribution from the eye source, that is, the
lens. Since the lens does not exist in the scene, as of now [TODO - this might
change later], there is no $W_e^0$. While this should be implemented in the
future, a hidden benefit of the current model is that there are no special cases
for paths where $t = 0$, which is impossible to handle on the GPU as any such
path could contribute to any arbitrary pixel [the light path is, after all,
directly hitting the lens]. So this has now become:
  \begin{align}
    \alpha_i^E = \Sigma_{i\leq 0} 
  \end{align}
  The amount that each path contributes to F differ greatly, so naively weighing
each contribution uniformly by something like 1.0f, or even, by inverse path
length, will producy very noisy images, and would not be worth the additional
computation time that BDPT requires, as shown below. To calculate the PDF of In
order to get good samples from the contributions, they must be combined in an
optimal manner using Multiple Importance Sampling: each contribution is weighted
by the PDF of the entire path itself:
  \begin{align*}
    F &\equiv \Sigma_s\Sigma_tW_{s, t}C_{s, t}\\
    C_{s, t} &= \alpha^L_s c_{s, t} \alpha^E_t
  \end{align*}
    where $W_{s, t}$ is the MIS weight of path $\bar{y} \leftarrow \bar{z}$, and
$C_{s, t}$ is the unweighted contribution of the same path. For clarity,
$C^U_{s, t}$ as described below is the unweighted contribution of the
    algorithm:
  \begin{align}
    C^u_{s, t} = (\alpha^L_s c_{s, t} \alpha^E_t) \equiv
    \frac{f_j(\bar{v})}{p(\bar{v})}
  \end{align}
  thus
  \begin{align}
    F = \Sigma_s\Sigma_tW_{s, t}C^u_{s, t}
  \end{align}
    The formula for Multiple Importance Sampling on BDPT is as so:
  \begin{align}
    MIS_{st}(\bar{x}) = \frac{P_s(\bar{x})}{\Sigma_i{P_i(\bar{x})}} 
  \end{align}
    Conceptually, the numerator $P_s(\bar{x})$ is the path density that was
generated, while the denominator is the path density of other connection
strategies that could could have, in theory,$^{[1]}$ created the path. The
straight forward implementation of this has a lot of problems gone unresolved in
Veach's thesis; PBR 3rd Ed, chapter 16 pg 1014-1016 addresses and describe these
problems. Primarily, $P_i(\bar{x})$ will overflow CLFloat (due to the distance
in the Geometric term), and the straight forward implementation is very long,
hard to debug, and has a poor time complexity. But the important part is the end
result of their solution:
  \begin{align}
    r_i (\bar{x}) = 
    \begin{cases}
      1.0f, & i = s,\\
      \frac{\overleftarrow{P}(\bar{x_i})}{\overrightarrow{P}(\bar{x_i})} *
      r_{i+1}(\bar{x}), & i < s,\\
      \frac{\overrightarrow{P}(\bar{x_{i-1}})}{\overleftarrow{P}(\bar{x_{i-1}})}
      * r_{i-1}(\bar{x}), & i > s
    \end{cases}
  \end{align}
    However, DTOADQ can simplify a bit further; unlike above, there are two
paths for which to evaluate. Specifically, in the above term, for where $i < s$,
we have $\frac{\overleftarrow{p}(\bar{x}_i)}{\overrightarrow{p}(\bar{x}_i)}$,
which is the light path (i is iterating towards light-length s). The eye
contribution is inversed from this, (also iterating towards s, hence why
$\bar{x}_{i-1}$ is used). Another way of saying it, is $\bar{x} = \bar{y} ||
\bar{z}_{rev}$. In DTOADQ's case, the weights are calculated in the same
direction relative to its originator; in this case the perspective of the eye
path.
  \begin{align}
    \mathcal{W}_i(\bar{x}) \equiv r_i(\bar{x})\\
    \mathcal{W}_i (\bar{x}) = 
    \begin{cases}
      1.0f, & i = 0,\\
      \frac{\overleftarrow{p_i}(\bar{x_i})}{\overrightarrow{p_i}(\bar{x_i})} *
      \mathcal{W}_{i-1}(\bar{x}), & i > 0
    \end{cases}
  \end{align}
    And then the expanded form of $\mathcal{W}_i(\bar{x})$ looks like
  \begin{align}
    \mathcal{W}_i(\bar{x}) = \Pi_{n=2}^i(\frac{\overleftarrow
    {p_{\sigma}}(x_n)\overleftarrow {G}(x_n)}
    {\overrightarrow{p_{\sigma}}(x_n)\overrightarrow{G}(x_n)})
    * \frac{\overleftarrow{p_A}(x_1)}{\overrightarrow{p_A}(x_1)}
  \end{align}
    And now the MIS looks like:
  \begin{align}
    MIS_{s, t}(\bar{y}, \bar{z}) =
    \frac{1.0f}{\mathcal{W}_s(\bar{y}) + \mathcal{W}_t(\bar{z}) + 1.0f}
  \end{align}
    Thus, the final rendering equation for DTOADQ looks like such:
  \begin{align}
    P = \frac{1.0f}{N} \Sigma_{0}^{N} \Sigma_{s \le 0} \Sigma_{t \le 0}
    \alpha^L_s c_{s,t} \alpha^E_t MIS_{s, t}(\bar{y}, \bar{z})
  \end{align}

@@

@@

  [Henyey]
   Diffuse Radation in the Galaxy, 1942. L, G. Henyey and J. L. Greenstein. 
\end{document}
\message{ !name(test.tex) !offset(-707) }

>>>>

  In order to compute a normal of an SDF, you must perform a gradient
approximation: the central distance onf the SDF at P. For example, the
difference quotient:
  \[\frac{(f(x+h) - f(x))}{h}\]
  The most obvious way to calculate this is the six-point gradient; where for
each axis you compute $f(x+h) - f(x-h)$, and normalize the result. This
returns a vector pointing in the direction where the map's SDF changes the
most: the normal. A four point gradient is possible; as long as each each
point of the gradient is multiplied by $\epsilon$ before normalizing. The
speedup from six to four is worthwhile, as computing each gradient is
expensive as it involves a sdf map call. There are more utilities with
gradients not yet currently provided in DTOADQ but are worth mentioning, a
few are: an additional normal call using an alternative (higher-quality) SDF
map can be used to compute a 'normal' bumpmap, and analyzing the gradient by
hand allows a cheap normal computation of primitives such as a spheres.

  Overall, it seems that while SDFs are fun to play around with, for the case of
  writing a generalized light transport algorithm to apply to SDFs causes a lot
  of problems while being generally limiting. The biggest problem is dealing
  with transparent materials, and dealing with visibility checks. For
  transparent materials, it's required that the ray march 'through' the SDF
  primitive, and find the exit. The solution is to apply a 'shell' to the SDF,
  or in other words, force the SDF to be unsigned:

\begin{lstlisting}
float Shell ( float dist, float r ) {
  return fabs(dist) - r*0.5f;
}
\end{lstlisting}

  \section{OpenCL}
	OpenCL, and GPU programming in general, is where I am least knowledgeable or
interested. Well, OpenCL is easy enough to work with compared to OGL or CUDA,
however I'm not good at dealing with the intrinsics of the GPU: optimizing code,
knowing how memory and thread/workers should be handled, etc. For example,
probably the biggest problem with the DTQ kernel right now is it is a single
file megakernel. There just was never any interest to split it into multiple
files nor multiple kernels. However, working directly with the OpenCL api is
easy enough, and I did write my own OpenCL module
to handle OCL in a rather convenient method, for example, the DTOADQ kernel call
looks as such:
	\begin{lstlisting}
	shared_info.image.Lock(); // Acquire OCL Mem from OGL image object
	ocl.Run(
	  ocl.CLStoreMem(shared_info.rw_image),
	  shared_info.image.RWrite(),
	  ocl.CLStoreMem(shared_info.image_metadata),
	  shared_info.camera,
	  shared_info.timer,
	  ocl.CLPredefinedMem(imgs),
	  ocl.CLStoreMem(shared_info.ocl_material),
	  debug_values,
	  ocl.CLStoreMem(shared_info.rng_states),
	  // kernel size
	  shared_info.image.x, shared_info.image.y
	);
	shared_info.image.Unlock();

@@

  The connection strategy below describes how to connect the edges of the two
paths. There is no $t = 0$ case as no physical camera lens exists on the scene.
  \begin{align}
    c_{s, t} =
    \begin{cases}
      L_e(L_0 \rightarrow L_1), &s = 0, t > 1,\\
      f_s(L_0 \rightarrow L_1 \rightarrow L2) G(E_1 \leftrightarrow
      L_1) f_s(L_1 \rightarrow E_1 \rightarrow E_0), &s > 0, t > 1
    \end{cases}
  \end{align}
  In code:
  \begin{lstlisting}
Spectrum contribution = E1.irradiance * L1->irradiance;

@@

    Another question to consider is if the path should be able to reflect on the
light source. While under real circumstances, this is allowed, the contribution
is negligible and dismissed in DTOADQ. Also, there are no cases for $s = s, t =
0|1$ as the lens does not have a physical existence in the SDF map. The $t = 0$
case, where a photon from the light source hits the lens directly, is rather
pointless anyways, as a path generated by light-tracing would have to be able to
contribute to any pixel, which is not feasible on a GPU.
\\

The amount that each path contributes differ greatly, so naively weighing
each contribution uniformly by something like 1.0f, or even, by inverse path
length, will producy very noisy images, and would not be worth the additional
computation time that BDPT requires. Thus, as mentioned earlier, probability
distributions of the entire path are taken into account in order to generate
multiple importance samples.
    The formula for multiple importance sampling on BDPT is as so:
  \begin{align}
    MIS_{st}(\bar{x}) = \frac{P_s(\bar{x})}{\Sigma_i{P_i(\bar{x})}} 
  \end{align}
    Conceptually, the numerator $P_s(\bar{x})$ is the path density that was
generated, while the denominator is the path density of other connection
strategies that could could have, in theory, created the path. The
straight forward implementation of this has a lot of problems gone unresolved in
Veach's thesis; PBR 3rd Ed, chapter 16 pg 1014-1016 addresses and describe these
problems. Primarily, $P_i(\bar{x})$ will overflow CLFloat (due to the distance
in the Geometric term), and the straight forward implementation is very long,
hard to debug, and has a poor time complexity. But the important part is the end
result of their solution:
  \begin{align}
    r_i (\bar{x}) = 
    \begin{cases}
      1.0f, & i = s,\\
      \frac{\overleftarrow{P}(\bar{x_i})}{\overrightarrow{P}(\bar{x_i})} *
      r_{i+1}(\bar{x}), & i < s,\\
      \frac{\overrightarrow{P}(\bar{x_{i-1}})}{\overleftarrow{P}(\bar{x_{i-1}})}
      * r_{i-1}(\bar{x}), & i > s
    \end{cases}
  \end{align}
    However, DTOADQ can simplify a bit further; unlike above, there are two
paths for which to evaluate. Specifically, in the above term, for where $i < s$,
we have $\frac{\overleftarrow{p}(\bar{x}_i)}{\overrightarrow{p}(\bar{x}_i)}$,
which is the light path (i is iterating towards light-length s). The eye
contribution is inversed from this, (also iterating towards s, hence why
$\bar{x}_{i-1}$ is used). . In DTOADQ's case, the weights are calculated in the
same direction relative to its originator; in this case the perspective of the
eye path. Not only this, but unlike any implementation I've found of BDPT, I
don't recalculate each path's probability and MIS every vertex. That is, usually
on calculating MIS for each connection, the MIS and probability values are
recalculated for each vertex in the path. As this is not possible given that
only the light-path exists in memory, an array for MIS/probability is stored
corresponding to each light vertex, and a single MIS/probability is stored
corresponding to $E_1$. Thus, the weight on each path looks like:
  \begin{align}
    \mathcal{W}_i(\bar{x}) \equiv r_i(\bar{x})\\
    \mathcal{W}_i (\bar{x}) = 
    \begin{cases}
      1.0f, & i = 0,\\
      \frac{\overleftarrow{p_i}(\bar{x_i})}{\overrightarrow{p_i}(\bar{x_i})} *
      \mathcal{W}_{i-1}(\bar{x}), & i > 0
    \end{cases}
  \end{align}
    And then the expanded form of $\mathcal{W}_i(\bar{x})$ looks like
  \begin{align}
    \mathcal{W}_i(\bar{x}) = \Pi_{n=2}^i(\frac{\overleftarrow
    {p_{\sigma}}(x_n)\overleftarrow {G}(x_n)}
    {\overrightarrow{p_{\sigma}}(x_n)\overrightarrow{G}(x_n)})
    * \frac{\overleftarrow{p_A}(x_1)}{\overrightarrow{p_A}(x_1)}
  \end{align}
    And now the MIS looks like:
  \begin{align}
    MIS_{s, t}(\bar{y}, \bar{z}) =
    \frac{1.0f}{\mathcal{W}_s(\bar{y}) + \mathcal{W}_t(\bar{z}) + 1.0f}
  \end{align}
    Thus, the final rendering equation for DTOADQ looks like such:
  \begin{align}
    P = \frac{1.0f}{N} \Sigma_{0}^{N} \Sigma_{s \le 0} \Sigma_{t \le 0}
    \alpha^L_s c_{s,t} \alpha^E_t MIS_{s, t}(\bar{y}, \bar{z})
  \end{align}

@@

  The camera handling function is based off glLookAt. The difference is that it
  supports field of view, depth of field and antialiasing. The model isn't
perfect and has a few bugs, mostly because the camera used in DTQ
is a pinhole camera, or as I call it delta camera. In general, very little time
was spent on the camera mechanics. In the future, there will be a physical
camera lens in the SDF, so as to actually simulate depth of field, aperture,
etc. Shown below is an exaggerated depth-of-field in the raytracer.
\\

  \includegraphics[scale=0.5]{rt_dof}

@@

  [Henyey]
   Diffuse Radation in the Galaxy, 1942. L, G. Henyey and J. L. Greenstein. 
\end{document}
\message{ !name(test.tex) !offset(-703) }
